
---
title: "Interface with other single-cell analysis toolkits"
author: "Suoqin Jin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mainfont: Arial
vignette: >
  %\VignetteIndexEntry{RNA velocity analysis of cellular dynamics using nlvelo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```


This vignette shows how to extract the required data for nlvelo from other existing single-cell analysis toolkits, including [Seurat](https://satijalab.org/seurat/). 


## Extract the following data information for velocity analysis
- Cell embedding for visualization
- Cell labels
- Cell-cell distance matrix


## Extract the required data from a Seurat V3 object
The required data can be obtained from the Seurat object by
```{r}
# Cell embedding for visualization
emb = Embeddings(seurat_object, "umap")
# Cell labels
labels <- Idents(seurat_object) # the cell labels
# Cell-cell distance matrix
cell.dist <- as.dist(1-armaCor(t(Embeddings(seurat_object, "pca"))))
```

## Extract the required data from a Seurat V2 object
The required data can be obtained from the Seurat object by
```{r}
# Cell embedding for visualization
emb <- seurat_object@dr$umap@cell.embeddings
# Cell labels
labels <- seurat_object@idents # the cell labels
# Cell-cell distance matrix
cell.dist <- as.dist(1-armaCor(t(seurat_object@dr$pca@cell.embeddings)))

```
