
---
title: "RNA velocity analysis of cellular dynamics using nlvelo"
author: "Suoqin Jin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mainfont: Arial
vignette: >
  %\VignetteIndexEntry{RNA velocity analysis of cellular dynamics using nlvelo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  root.dir = './'
)
```


This vignette outlines the steps of RNA velocity analysis of cellular dynamics using nlvelo by applying it to studying the interfollicular epidermal cell transition dynamics during skin homeostasis. 


## Load the required libraries
```{r message=FALSE,warning=FALSE}
library(nlvelo)
library(igraph)
```

## Load loom files
```{r}
# Here we load loom files of two replicates and then combined them together
ldat1 <- read.loom.matrices("/Users/suoqinjin/Dropbox/example_nlvelo/ms_uw_1.loom")
ldat2 <- read.loom.matrices("/Users/suoqinjin/Dropbox/example_nlvelo/ms_uw_2.loom")
ldat <- ldat1
emat1 <- ldat1$spliced;
emat2 <- ldat2$spliced;
emat <- cbind(emat1, emat2);
nmat1 <- ldat1$unspliced;
nmat2 <- ldat2$unspliced;
nmat <- cbind(nmat1, nmat2);
ldat$spliced <- emat; ldat$unspliced <- nmat;
```

## Take an existing embedding for visualization
```{r}
ydata <- read.delim("/Users/suoqinjin/Dropbox/example_nlvelo/projectedData_ump_UW_IFE.txt", row.names = 1, header = T)
emb <- as.matrix(ydata)
```

## Subset the loom files for interested cells
```{r}
# processing cell barcodes is data-specific
barcodes_loom <- colnames(ldat$spliced)
barcodes_loom <- substr(barcodes_loom,15,35) # change it if necessary
barcodes_loom <- gsub(":", "_", barcodes_loom) # change it if necessary
barcodes <- row.names(ydata)
barcodes <- substr(barcodes,1,21) # change it if necessary
# subset the loom files by matching the cell barcodes with the loaded cell embedding
idx = match(barcodes,barcodes_loom)
ldat$spliced <- ldat$spliced[,idx]; ldat$unspliced <- ldat$unspliced[,idx];
colnames(ldat$spliced) <- row.names(ydata);colnames(ldat$unspliced) <- row.names(ydata);
emat <- ldat$spliced; nmat <- ldat$unspliced;
```

## Load cell labels and define cell colors
```{r}
# cell.label should be a named factor
cell.label <- read.table("/Users/suoqinjin/Dropbox/example_nlvelo/clusters_UW_IFE.txt",sep = '\t',row.names=1,header=T)$cluster
names(cell.label) <- rownames(ydata)
# cell.colors should be also named
color.use <- c('#FF1622','#00b938','#3657df','#603A91','#F29233','#D6B239')
names(color.use) <- levels(cell.label)
cell.colors <- color.use[cell.label]
names(cell.colors) <- rownames(ydata)
```

## Compute the cell-cell distance
```{r, fig.show = 'hide', results='hide', message=FALSE, warning=FALSE}
# Pagoda2 is used to generate cell-cell distance matrix. You can alternatively generate those using other tools, such as Seurat
library(pagoda2)
row.names(emat) <- make.unique(row.names(emat),sep = '_')
r <- Pagoda2$new(emat,modelType='plain',trim=10,log.scale=T)
r$adjustVariance(plot=T,do.par=T,gam.k=10)
r$calculatePcaReduction(nPcs=100,n.odgenes=3e3,maxit=300)
# Calculate cell-cell distance matrix
cell.dist <- as.dist(1-armaCor(t(r$reductions$PCA)))
```

## Filter genes based on the minimum average expresion count (in at least one of the clusters)
```{r}
emat <- filter.genes.by.cluster.expression(emat,cell.label,min.max.cluster.average = 0.1) # filtered emat matrix
nmat <- filter.genes.by.cluster.expression(nmat,cell.label,min.max.cluster.average = 0.05)
length(intersect(rownames(emat),rownames(nmat)))
```

## Estimate RNA velocity using nlvelo (nonlinear model)
```{r}
# using gene-relative model with k=30 cell kNN pooling and using top/bottom 2% quantiles for gamma fit:
rvel <- gene.relative.velocity.estimates(emat,nmat,kCells=30,cell.dist=cell.dist,fit.quantile=0.02)
```

## Visualize velocity on an existing embedding
```{r, fig.width=4,fig.height = 4,  fig.wide = TRUE, fig.align = "center"}
show.velocity.on.embedding.cor(emb,rvel,n=500,scale='sqrt',cell.colors=ac(cell.colors,alpha=0.8),cex=0.8,arrow.scale=3,show.grid.flow=TRUE,min.grid.cell.mass=0.5,grid.n=40,arrow.lwd=0.5,do.par=T,cell.border.alpha = 0.001,
                               xlab = "UMAP1", ylab = "UMAP2")
```

### Save the workspace
```{r}
save(emb, rvel, emat, nmat, cell.colors, cell.dist, file = "nlvelo_analysis_IFE.RData")
```

